---
title: "p8105_hw6_jy3360"
author: "Candice Yu"
date: "2023-11-20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Problem 1
### Part a
The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.

* Create a `city_state` variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. 
* Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. 
* Also omit Tulsa, AL – this is a data entry mistake. 
* For this problem, limit your analysis those for whom victim_race is white or black. 
* Be sure that victim_age is numeric.

```{r}
homicide_data <- read_csv("homicide-data.csv") %>%
  janitor::clean_names() %>%
  mutate(city_state = paste(city, state, sep = ", "),
         solved = as_factor(if_else(disposition %in% c("Closed without arrest", "Closed by arrest"), 1, 0)),
         victim_age = as.numeric(victim_age)) %>%
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
         victim_race %in% c("White", "Black"))
```

### Part b
For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. 

* Save the output of glm as an R object; 
* apply the broom::tidy to this object; 
* and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
# data for Baltimore, MD
baltimore_data <- homicide_data %>%
  filter(city_state == "Baltimore, MD")

# logistic regression for Baltimore
baltimore_glm <- glm(solved ~ victim_age + victim_sex + victim_race, 
                     family = binomial, 
                     data = baltimore_data)

# save the model output and tidy it
baltimore_model_tidied <- broom::tidy(baltimore_glm)

# extract the estimate and confidence interval for male vs female victims
baltimore_coefs <- coef(summary(baltimore_glm))

male_female_or <- exp(baltimore_coefs["victim_sexMale", "Estimate"])
male_female_or_ci <- exp(confint(baltimore_glm)["victim_sexMale",])

```

The estimate of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed is `r male_female_or` and the confidence interval is (`r male_female_or_ci`).

### Part c
Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. 

Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
# Assuming your data is in a dataframe called cleaned_data

# Define a function to perform glm and extract the necessary information
run_glm <- function(df) {
  glm_model = glm(solved ~ victim_age + victim_sex + victim_race, family = binomial, data = df)
  tidy_glm = broom::tidy(glm_model)
  confint_glm = confint(glm_model)
  
  # Extract the estimate, std. error, and CI for victim_sexMale
  tidy_glm %>%
    filter(term == "victim_sexMale") %>%
    mutate(
      or = exp(estimate),
      lower_ci = exp(confint_glm["victim_sexMale", "2.5 %"]),
      upper_ci = exp(confint_glm["victim_sexMale", "97.5 %"])
    )
}

# Apply the function to each subset of data for each city
city_models <- homicide_data %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(model = purrr::map(data, run_glm)) %>%
  select(-data) %>%
  unnest(model)

```

### Part d
Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r, fig.width = 8, fig.height = 10}
# Create a plot of the odds ratios
ggplot(city_models, aes(x = reorder(city_state, or), y = or)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.25) +
  coord_flip() +  # flip the axes 
  xlab("City") +
  ylab("Odds Ratio") +
  ggtitle("Adjusted Odds Ratios with CI for Solving Homicides (Male vs Female Victims)")

```

**Comments**

- The ORs for most cities are less than 1, suggesting that, after keeping all other variables fixed, homicides with female victims are more likely to be solved than those with male victims.
- The cities at the top of the plot, such as Fresno, CA, and Minneapolis, MN, have ORs greater than 1, indicating a relatively higher likelihood of solving cases with male victims in these cities.
- Some cities, like Tulsa, OK, and Nashville, TN, have confidence intervals that cross the OR of 1, suggesting that there is no statistically significant difference in the likelihood of solving homicides between male and female victims in those cities.
- The confidence intervals for cities like Fresno, CA, do not cross 1, suggesting a statistically significant difference in the likelihood of solving homicides with male victims compared to female victims.


# Problem 2
For this problem, we’ll use the Central Park weather data similar to data we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

The boostrap is helpful when you’d like to perform inference for a parameter/ value/ summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with `tmax` as the response with `tmin` and `prcp` as the predictors, and are interested in the distribution of two quantities estimated from these data:

* \( \hat{r}^2 \) 
* \( \log(\hat{\beta}_1 \times \hat{\beta}_2) \)

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. 

- Plot the distribution of your estimates, and describe these in words. 
- Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for \( \hat{r}^2 \) and \( \log(\hat{\beta}_1 \times \hat{\beta}_2) \). 

Note: broom::glance() is helpful for extracting \( \hat{r}^2 \) from a fitted regression, and `broom::tidy()` (with some additional wrangling) should help in computing \( \log(\hat{\beta}_1 \times \hat{\beta}_2) \)). 

```{r}
# define a function to perform the linear regression and extract r^2 and beta_product = log(beta1 * beta2)
analyze_sample <- function(sample_data) {
  model <- lm(tmax ~ tmin + prcp, data = sample_data)
  glance_data <- broom::glance(model)
  tidy_data <- broom::tidy(model)
  r_squared <- glance_data$r.squared
  beta_product <- exp(log(tidy_data$estimate[2]) + log(tidy_data$estimate[3]))
  return(c(r_squared, beta_product))
}

# Generate the bootstrap samples using rerun and analyze each sample
set.seed(123) 
bootstrap_samples <- rerun(5000, sample_n(weather_df, nrow(weather_df), replace = TRUE))
bootstrap_results <- map(bootstrap_samples, analyze_sample)

# Extract r^2 and log(beta1 * beta2) estimates
r_squared <- map_dbl(bootstrap_results, 1)
log_beta_product <- map_dbl(bootstrap_results, 2)

```
```{r}
# Plot the distribution of r^2 estimates
ggplot(data.frame(r_squared), aes(x = r_squared)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  labs(title = "Distribution of R-squared Estimates", x = "R-squared", y = "Frequency")

# Plot the distribution of log(beta1 * beta2) estimates
ggplot(data.frame(log_beta_product), aes(x = log_beta_product)) +
  geom_histogram(bins = 30, fill = "red", color = "black") +
  labs(title = "Distribution of log(beta1 * beta2) Estimates", x = "log(beta1 * beta2)", y = "Frequency")

# Calculate the 95% confidence intervals
r_squared_ci <- quantile(r_squared, probs = c(0.025, 0.975))
log_beta_product_ci <- quantile(log_beta_product, probs = c(0.025, 0.975), na.rm = TRUE)

```

- 95% CI for \( \hat{r}^2 \) is [`r r_squared_ci[1]`, `r r_squared_ci[2]`]
- 95% CI for \( \log(\hat{\beta}_1 \times \hat{\beta}_2) \) is [`r log_beta_product_ci[1]`, `r log_beta_product_ci[2]`]. 
